First Order Alternation

### 2019-02-07 quantifiers/body.tex
ok
201902070143

The first order alternating automaton recognizing the (encoding of
the) $\clang{B}$ language is simply obtained by negating the theory
atoms $t \geq u$ and $u - y \neq 1$ and interchanging conjunctions
with disjunctions in the rules of the alternating automaton encoding
$B$, denoted as $\overline{B}$ in Fig. \ref{fig:ta} (bottom
right). Finally, we check the emptiness of the alternating automaton
consisting of the rules of $A$ and $\overline{B}$, started in the
initial configuration $s_1(0,0) \wedge q_1(0,0)$. Our lazy annotation
semi-algorithm finds that this automaton is empty in \comment[ri]{...}
seconds.

@@

\begin{theorem}[\cite{BorgerGraedelGurevich97}]\label{thm:skolem}
  Given $Q_1 x_1 \ldots Q_n x_n ~.~ \phi$ a first order sentence,
  where $Q_1, \ldots, Q_n \in \set{\exists,\forall}$ and $\phi$ is
  quantifier-free, let $\eta_i \isdef f_i(y_1,\ldots,y_{k_i})$ if $Q_i
  = \forall$ and $\eta_i \isdef x_i$ if $Q_i = \exists$, where $f_i$
  is a fresh function symbol and $\set{y_1, \ldots, y_{k_i}} =
  \set{x_j \mid j < i,~ Q_j = \exists}$. Then the entailment \(Q_1 x_1
  \ldots Q_n x_n ~.~ \phi \models \phi[\eta_1/x_1,\ldots,\eta_n/x_n]\)
  holds.
\end{theorem}
\paragraph{Proof.} See \cite[Theorem 2.1.8]{BorgerGraedelGurevich97} and
  \cite[Lemma 2.1.9]{BorgerGraedelGurevich97}. \qed

@@

\begin{table}[htb]
\vspace*{-\baselineskip}
\begin{center}
{\fontsize{7}{8}\selectfont
\begin{tabular}{||l|c|c|c|c|c||}
\hline
Example & $\len{\A}$ (bytes) & $L(\A)=\emptyset$ ? & Nodes Expanded & Nodes Visited & Time (ms) \\
\hline
incdec.pa & 499 & no & 21 & 17 & 779 \\
\hline
localdec.pa & 678 & no & 49 & 35 & 1814 \\
\hline
ticket.pa & 4250 & no & 229 & 91 & 9543 \\
\hline
count\_thread0.pa & 9767 & no & 154 & 128 & 8553 \\
\hline
count\_thread1.pa & 10925 & no & 766 & 692 & 76771 \\
\hline
local0.pa & 10595 & no & 73 & 27 & 1431 \\
\hline
local1.pa & 11385 & no & 1135 & 858 & 101042 \\
\hline
array\_rotation.ada & 1834 & yes & 9 & 8 & 1543 \\
\hline
array\_simple.ada & 3440 & yes & 11 & 10 & 6787 \\
\hline
array\_shift.ada & 874 & yes & 6 & 5 & 413 \\
\hline
abp.ada & 6909 & no & 52 & 47 & 4788 \\
\hline
train.ada & 1823 & yes & 68 & 67 & 7319 \\
\hline
hw1.ada & 322 & Solver Error & / & / & / \\
\hline
hw2.ada & 674 & yes & 20 & 22 & 4974 \\
\hline
rr-crossing.foada & 1780 & yes & 67 & 67 & 7574 \\
% \hline
% running2.foada & 1217 & yes & 13 & 13 & 788 \\
\hline
train-simple1.foada & 5421 & yes & 43 & 44 & 2893 \\
\hline
train-simple2.foada & 10177 & yes & 111 & 113 & 8386 \\
\hline
train-simple3.foada & 15961 & yes & 196 & 200 & 15041 \\
% \hline
% fischer-mutex1.foada & 1566 & no & 31 & 31 & 1552 \\
\hline
fischer-mutex2.foada & 3000 & yes & 23 & 23 & 808 \\
\hline
fischer-mutex3.foada & 4452 & yes & 33 & 33 & 1154 \\
\hline
\end{tabular}
}
\caption{Experiments with First Order Alternating Automata}
\label{tab:experiments}
\end{center}
\vspace*{-3\baselineskip}
\end{table}

>>>>

The first order alternating automaton recognizing the (encoding of
the) $\clang{B}$ language is simply obtained by negating the theory
atoms $t \geq u$ and $u - y \neq 1$ and interchanging conjunctions
with disjunctions in the rules of the alternating automaton encoding
$B$, denoted as $\overline{B}$ in Fig. \ref{fig:ta} (bottom
right). Finally, we check the emptiness of the alternating automaton
consisting of the rules of $A$ and $\overline{B}$, started in the
initial configuration $s_1(0,0) \wedge q_1(0,0)$. Our lazy annotation
semi-algorithm finds that this automaton is empty in 347 ms
seconds.

@@

\begin{theorem}[\cite{BorgerGraedelGurevich97}]\label{thm:skolem}
  Given $Q_1 x_1 \ldots Q_n x_n ~.~ \phi$ a first order sentence,
  where $Q_1, \ldots, Q_n \in \set{\exists,\forall}$ and $\phi$ is
  quantifier-free, let $\eta_i \isdef f_i(y_1,\ldots,y_{k_i})$ if $Q_i
  = \forall$ and $\eta_i \isdef x_i$ if $Q_i = \exists$, where $f_i$
  is a fresh function symbol and $\set{y_1, \ldots, y_{k_i}} =
  \set{x_j \mid j < i,~ Q_j = \exists}$. Then the entailment \(Q_1 x_1
  \ldots Q_n x_n ~.~ \phi \models \phi[\eta_1/x_1,\ldots,\eta_n/x_n]\)
  holds.
\end{theorem}
\paragraph{\em Proof.} See \cite[Theorem 2.1.8]{BorgerGraedelGurevich97} and
  \cite[Lemma 2.1.9]{BorgerGraedelGurevich97}. \qed

@@

\begin{table}[htb]
%\vspace*{-\baselineskip}
\begin{center}
{\fontsize{7}{8}\selectfont
\begin{tabular}{||l|c|c|c|c|c|c|c|c||}
\hline
Example & $\len{\A}$ (bytes) & Nb of Predicates & Nb of Variables & Nb of Transitions & $L(\A)=\emptyset$ ? & Nodes Expanded & Nodes Visited & Time (ms) \\
\hline
incdec.pa & 499 & 3 & 1 & 12 & no & 21 & 17 & 779 \\
\hline
localdec.pa & 678 & 4 & 1 & 16 & no & 49 & 35 & 1814 \\
\hline
ticket.pa & 4250 & 13 & 1 & 73 & no & 229 & 91 & 9543 \\
\hline
count\_thread0.pa & 9767 & 14 & 1 & 126 & no & 154 & 128 & 8553 \\
\hline
count\_thread1.pa & 10925 & 15 & 1 & 135 & no & 766 & 692 & 76771 \\
\hline
local0.pa & 10595 & 13 & 1 & 117 & no & 73 & 27 & 1431 \\
\hline
local1.pa & 11385 & 14 & 1 & 126 & no & 1135 & 858 & 101042 \\
\hline
array\_rotation.ada & 1834 & 8 & 7 & 7 & yes & 9 & 8 & 1543 \\
\hline
array\_simple.ada & 3440 & 9 & 5 & 8 & yes & 11 & 10 & 6787 \\
\hline
array\_shift.ada & 874 & 6 & 5 & 5 & yes & 6 & 5 & 413 \\
\hline
abp.ada & 6909 & 16 & 14 & 28 & no & 52 & 47 & 4788 \\
\hline
train.ada & 1823 & 10 & 4 & 26 & yes & 68 & 67 & 7319 \\
\hline
hw1.ada & 322 & 3 & 2 & 5 & Solver Error & / & / & / \\
\hline
hw2.ada & 674 & 7 & 2 & 8 & yes & 20 & 22 & 4974 \\
\hline
rr-crossing.foada & 1780 & 10 & 1 & 16 & yes & 67 & 67 & 7574 \\
% \hline
% running2.foada & 1217 & yes & 13 & 13 & 788 \\
\hline
train-simple1.foada & 5421 & 13 & 1 & 61 & yes & 43 & 44 & 2893 \\
\hline
train-simple2.foada & 10177 & 16 & 1 & 118 & yes & 111 & 113 & 8386 \\
\hline
train-simple3.foada & 15961 & 19 & 1 & 193 & yes & 196 & 200 & 15041 \\
% \hline
% fischer-mutex1.foada & 1566 & no & 31 & 31 & 1552 \\
\hline
fischer-mutex2.foada & 3000 & 11 & 2 & 23 & yes & 23 & 23 & 808 \\
\hline
fischer-mutex3.foada & 4452 & 16 & 2 & 34 & yes & 33 & 33 & 1154 \\
\hline
\end{tabular}
}
\caption{Experiments with First Order Alternating Automata}
\label{tab:experiments}
\end{center}
%\vspace*{-3\baselineskip}
\end{table}
